Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-10-31 thesis_chapter_3.Rmd
new chapter 3 files

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
output:
  word_document:
    reference_docx: markdown/wiley_template.docx
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
bibliography: references.bib
csl: markdown/limnologyandoceanography.csl
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'. Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain. The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
---

@@

library(ggplot2)
#library(rgdal)
library(gridExtra)
library(ggpubr)
library(mapproj)
library(vegan)
library(dplyr)
library(ggpubr)
library(maps)
library(flextable)
library(officer)
library(tidync)
library(tidyverse)
library(hexbin)
library(lubridate)

@@

A range of natural and anthropogenic objects floats in the ocean, and macroalgae are regarded as one of the most important of these passive dispersal mechanisms of marine taxa. Floating kelps act as passive dispersal mechanisms for a range of marine taxa and is sometimes referred to as the ‘tumble-weed of the ocean’ [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Kelps also function as input of allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats [@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].

Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant. The positively buoyant pneumatocysts also cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts varies among species. For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases, the stipe itself may be air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating species of Chlorophyta and Rhodophyta, Phaeophyceae species are the most reported forms of floating algae. This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognized as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterey Bay. The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear. In addition, the trajectories of tagged kelp were more consistent with the formation of eddys during winter. Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to the raft. The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate. Genetic analyses showed these species can raft up to $\sim$ 390 km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017]. Work by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

@@

#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final.png")))
ggarrange(map_chp3_combined)

@@

```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E. maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}

@@

```{r data wrangling, message=FALSE, warning=FALSE, include=FALSE}

# Read in simulation data and data wrangling

p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
p_RK4_mean <- p_RK4 %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
p_RK4_mean$type <- "1" # add extra column 'type' to dataframe and assigned 'RK4' label

Min_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H100_W00.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H100_W00_mean <- Min_Kfloat_H100_W00 %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H100_W00_mean$type <- "2" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H095_W05.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H095_W05_mean <- Min_Kfloat_H095_W05 %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H095_W05_mean$type <- "3" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H090_W10.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H090_W10_mean <- Min_Kfloat_H090_W10 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H090_W10_mean$type <- "4" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Min_KFloat_H085_W15.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H085_W15_mean <- Min_Kfloat_H085_W15 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H085_W15_mean$type <- "5" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_Kfloat_H100_W00_mean <- Mean_Kfloat_H100_W00 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_Kfloat_H100_W00_mean$type <- "6" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H095_W05.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H095_W05_mean <- Mean_KFloat_H095_W05 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H095_W05_mean$type <- "7" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H090_W10.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H090_W10_mean <- Mean_KFloat_H090_W10 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H090_W10_mean$type <- "8" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H085_W15.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Mean_KFloat_H085_W15_mean <- Mean_KFloat_H085_W15 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H085_W15_mean$type <- "9" # add extra column 'type' to dataframe and assigned 'kelp' label

Max_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H100_W00.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H100_W00_mean <- Max_Kfloat_H100_W00 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H100_W00_mean$type <- "10" # add extra column 'type' to dataframe and assigned 'kelp' label

Max_Kfloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H095_W05.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H095_W05_mean <- Max_Kfloat_H095_W05 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H095_W05_mean$type <- "11" # add extra column 'type' to dataframe and assigned 'kelp' label

Max_Kfloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H090_W10.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H090_W10_mean <- Max_Kfloat_H090_W10 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H090_W10_mean$type <- "12" # add extra column 'type' to dataframe and assigned 'kelp' label

Max_Kfloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Max_KFloat_H085_W15.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Max_Kfloat_H085_W15_mean <- Max_Kfloat_H085_W15 %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H085_W15_mean$type <- "13" # add extra column 'type' to dataframe and assigned 'kelp' label

total_data <- rbind(Min_Kfloat_H100_W00 , Min_Kfloat_H095_W05, Min_Kfloat_H090_W10, Min_Kfloat_H085_W15, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15, Max_Kfloat_H100_W00, Max_Kfloat_H095_W05, Max_Kfloat_H090_W10, Max_Kfloat_H085_W15)

## April

p_RK4_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/RK4_PFloat_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00")) 
p_RK4_mean_April <- p_RK4_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

p_RK4_April$type <- "RK4_April" # add extra column 'type' to dataframe and assigned 'RK4' label
p_RK4_mean_April$type <- "14" # add extra column 'type' to dataframe and assigned 'RK4' label

Min_Kfloat_H100_W00_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H100_W00_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H100_W00_mean_April <- Min_Kfloat_H100_W00_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H100_W00_April$type <- "Min_Kfloat_H100_W00_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00_April$velocity <- with(Min_Kfloat_H100_W00_April, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H100_W00_mean_April$type <- "15" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H095_W05_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H095_W05_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H095_W05_mean_April <- Min_Kfloat_H095_W05_April %>% # calculate mean trajectory
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H095_W05_April$type <- "Min_Kfloat_H095_W05_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05_April$velocity <- with(Min_Kfloat_H095_W05_April, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H095_W05_mean_April$type <- "16" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H090_W10_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H090_W10_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H090_W10_mean_April <- Min_Kfloat_H090_W10_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

Min_Kfloat_H090_W10_April$type <- "Min_Kfloat_H090_W10_April" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10_April$velocity <- with(Min_Kfloat_H090_W10_April, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H090_W10_mean_April$type <- "17" # add extra column 'type' to dataframe and assigned 'kelp' label

Min_Kfloat_H085_W15_April <- tidync("OceanParcels/trajectory_data/chapter3_additional/Min_KFloat_H085_W15_April.nc") %>%
  hyper_tibble() %>%
  dplyr::select(trajectory, lon, lat, time, distance, kelpvelU, kelpvelV) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))
Min_Kfloat_H085_W15_mean_April <- Min_Kfloat_H085_W15_April %>% #calculate mean trajectory 
  group_by(t) %>% 
  summarise(across(lon:lat, mean, na.rm= TRUE))

>>>>

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[Department of Biodiversity and Conservation Biology, University of the Western Cape, Private Bag X17, Bellville 7535, South Africa], C Rautenbach^[National Institute of Water and Atmospheric Research, Hamilton: Gate 10 Silverdale Road Hillcrest, 3216], AJ Smit^[South African Environmental Observation Network, Elwandle Coastal Node, Port Elizabeth, South Africa]
output:
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  word_document:
    reference_docx: markdown/wiley_template.docx
bibliography: references.bib
csl: markdown/limnologyandoceanography.csl
abstract: "**Abstract**: Ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal vectors  of marine organisms and are regarded as the 'tumble-weed of the ocean'. Despite many studies that exist on the topic, the relative role of wind and surface currents in influencing the trajectory of macroalgal dispersal is still unknown. Past studies focus on kelp-rafts, which can vary greatly in size, making it difficult to form a consenses as to the relative role of wind versus surface currents. In addition, these studies have not considered surface-area in relation to drag characteristics of the macroalgae, both of which have been shown to play a role in the trajectory and accumulation of other flotsam. Advances in Lagrangian trajectory modeling have been rapid in recent years and allow the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagal trajectory and accumulation. We focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp 'particles' which incorporate drag with that of purely Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddys within the study domain. The results also point to the importance of sub-mesoscale features, such as eddys, as important surface accumulation zones for solitary floating macroalage."
---

@@

library(tidyverse)
library(flextable)
library(tidync)
library(ncdf4)
library(lubridate)
library(gridExtra)
library(ggpubr)
library(jpeg)
library(gridExtra)
library(grid)

@@

Passive dispersal mechanisms of marine taxa include a range of natural and anthropogenic objects that float in the ocean, and macroalgae are regarded as one of the most important of these. Floating kelps, sometimes referred to as the ‘tumble-weed of the ocean,’ are very important in this regard [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Kelps also function as allochthonous materials to near and distant ecosystems and in this manner support the ecological function in a range of adjacent habitats[@cartraud2021; @colombini2003; @Filbee-Dexter2018-wp; @zuercher2019; @Krumhansl2012; @vanderklift2008; @bustamante1995].

Some macroalgal species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant. The positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts vary among species. For example, species belonging to the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, and *Ecklonia maxima* have single, large pneumatocysts[@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases, the stipe itself may be air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating species of Chlorophyta and Rhodophyta, Phaeophyceae species are the most reported forms of floating algae. This is most likely because the green and red species reported to float are not actually positively buoyant, but instead are kept at the surface by gas trapped in-between or in their thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@tala2017; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @brooks2018; @wang2019] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgal trajectories being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown, although the role of wind has been recognized as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterey Bay. The results showed that kelp rafts, with little surface area exposed to the wind, were largely driven by a combination of wind and surface currents; however, the relative importance of wind and was not clear. In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter. Studies have also identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important mechanism for dispersal of floating bull kelp *Durvillaea antarctica* [@waters2008; @smith2002; @nikula2010]. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goose barnacles attached to the raft. The presence of goose barnacles suggests a long rafting time as these species have a slow growth rate. Genetic analyses showed these species can raft up to $\sim$ 390 km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic-form drag [@breivik2011; @griffin2017]. Work by the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

@@

#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_final_v2.png")))
ggarrange(map_chp3_combined)

@@

```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E. maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic. Schemetaic was adapted from @Dyer2018-rv."}

@@

```{r data readin, message=FALSE, warning=FALSE, include=FALSE}

total_summer <- read.csv(file = "data/chapter_3/total_summer.csv")
total_summer_mean <- read.csv(file = "data/chapter_3/total_summer_mean.csv")

total_autumn <- read.csv(file = "data/chapter_3/total_autumn.csv")
total_autumn_mean <- read.csv(file = "data/chapter_3/total_autumn_mean.csv")

total_winter <- read.csv(file = "data/chapter_3/total_winter.csv")
total_winter_mean <- read.csv(file = "data/chapter_3/total_winter_mean.csv")

total_spring <- read.csv(file = "data/chapter_3/total_spring.csv")
total_spring_mean <- read.csv(file = "data/chapter_3/total_spring_mean.csv")

total_passive_summer <- read.csv(file = "data/chapter_3/total_passive_summer.csv")
total_passive_summer_mean <- total_passive_summer %>% 
  group_by(t, type) %>%
  summarise(across(lon:lat, mean, na.rm= TRUE))
total_passive_autumn <- read.csv(file = "data/chapter_3/total_passive_autumn.csv")
total_passive_autumn_mean <- total_passive_autumn %>% 
  group_by(t, type) %>%
  summarise(across(lon:lat, mean, na.rm= TRUE))
total_passive_winter <- read.csv(file = "data/chapter_3/total_passive_winter.csv")
total_passive_winter_mean <- total_passive_winter %>% 
  group_by(t, type) %>%
  summarise(across(lon:lat, mean, na.rm= TRUE))
total_passive_spring <- read.csv(file = "data/chapter_3/total_passive_spring.csv")
total_passive_spring_mean <- total_passive_spring %>% 
  group_by(t, type) %>%
  summarise(across(lon:lat, mean, na.rm= TRUE))

total_eddy_hydro <- read.csv(file = "data/chapter_3/total_eddy_hydro.csv")
total_eddy_hydrostksdrag <- read.csv(file = "data/chapter_3/total_eddy_hydrostksdrag.csv")
total_eddy_hydrowinddrag <- read.csv(file = "data/chapter_3/total_eddy_hydrowinddrag.csv")
total_eddy_hydrowindstksdrag <- read.csv(file = "data/chapter_3/total_eddy_hydrowindstksdrag.csv")

total_eddy_passive_curr <- read.csv(file = "data/chapter_3/total_eddy_passive_curr.csv")
total_eddy_passive_currstks <- read.csv(file = "data/chapter_3/total_eddy_passive_currstks.csv")
total_eddy_passive_currwind <- read.csv(file = "data/chapter_3/total_eddy_passive_currwind.csv")
total_eddy_passive_currwindstks <- read.csv(file = "data/chapter_3/total_eddy_passive_currwindstks.csv")
